---
title: "Exploratory Analysis Flexdashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(tidyverse)

menustat_df = read_csv("data/menustat_combined.csv")
rest_locations = read_csv("data/restaurant_location.csv") 

final_cal_df = 
  menustat_df %>% 
  group_by(rest_type) %>% 
  filter(restaurant == "McDonald's" | restaurant == "Popeyes" | restaurant == "KFC" | restaurant == "Boston Market" | restaurant == "Chipotle" | restaurant == "White Castle" | restaurant == "Tim Hortons" | restaurant == "Little Caesers" | restaurant == "Wingstop" | restaurant == "Taco Bell" | restaurant == "Subway" | restaurant == "Wendy's" | restaurant == "Shake Shack" | restaurant == "Papa John's" | restaurant == "IHOP" | restaurant == "Pizza Hut" | restaurant == "Applebee's" | restaurant == "Panda Express" | restaurant == "Checkers" | restaurant == "Arby's" | restaurant == "Burger King" | restaurant == "Starbucks" | restaurant == "Panera Bread" | restaurant == "Tropical Smoothie Café" | restaurant == "Kriskpy Kreme" | restaurant == "Five Guys" | restaurant == "Buffalo Wild Wings" | restaurant == "Dominos" | restaurant == "Outback Steakhouse" | restaurant == "Moe's Southwest Grill" | restaurant == "Auntie Anne's" | restaurant == "Chick-Fil-A" | restaurant == "Sbarro" | restaurant == "Dunkin’ Donuts" | restaurant == "Hardee" | restaurant == "HOOTERS" | restaurant == "TGI FRIDAY'S"| restaurant == "JAMBA JUICE" | restaurant == "MARCO'S" | restaurant == "Red Lobster" | restaurant == "The Capital Grille" | restaurant == "Dave & Buster's" | restaurant == "Chuck E. Cheese's" | restaurant == "Perkins" | restaurant == "Jimmy Johns" | restaurant == "Baskin Robbins" | restaurant == "PF Chang's" | restaurant == "Dickey's Barbeque Pit" | restaurant == "The Cheesecake Factor" | restaurant == "California Pizza Kitchen" | restaurant == "Dairy Queen" | restaurant == "Qdoba" | restaurant == "Sonic"| restaurant == "Longhorn Steakhouse") %>%
  filter(quantile(calories, 0.95) > calories) 

distinct_food_categ =
  final_cal_df %>% 
  drop_na(rest_type) %>% 
   mutate(
    food_category = 
      recode(food_category, "Appetizers & sides" = "Appetizers & Sides")
    ) %>% 
  distinct(food_category)

rest_freq =
  rest_locations %>% 
  group_by(restaurant) %>% 
  mutate(restaurant = recode(restaurant, "APPLEBEES" = "APPLEBEE'S", "AUNTIE ANNES" = "AUNTIE ANNE'S PRETZELS", "Pizza Hut" = "PIZZA HUT", "POPEYE'S" = "POPEYES", "WingStop" = "WINGSTOP"))
```

Sidebar {.sidebar}
======================================================================

Initial Visualizations
======================================================================

### Distribution of Total Calories By Restaurant Type

```{r}
menustat_df %>% 
  group_by(rest_type) %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Items With Calorie Counts in Top 5% of Restaurants 

```{r}
menustat_df %>% 
  group_by(rest_type) %>% 
  filter(quantile(calories, 0.95) < calories) %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Frequency of Restaurants Across NYC

```{r}
rest_freq %>% 
  count(restaurant) %>% 
  mutate(neighbourhood = fct_reorder(restaurant, n)) %>% 
  plot_ly(x = ~restaurant, y = ~n, color = ~restaurant, type = "bar", colors = "viridis")
```


Calorie Distribution By Restaurant Type
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------


### Calorie Distribution in 2008

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2008") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calorie Distribution in 2012

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2012") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calorie Distribution in 2016

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2016") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calorie Distribution in 2020

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2020") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```


Fat Distribution By Restaurant Type
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------


### Trans Fat Distribution in 2008

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2008") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution in 2012

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2012") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution in 2016

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2016") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution in 2020

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2020") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```


Protein Distribution By Restaurant Type
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------


### Protein Distribution in 2008

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2008") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in 2012

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2012") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in 2016

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2016") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in 2020

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2020") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```


Nutritional Content by Food Category
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Calories Distribution by Food Category Overlayed

```{r}
final_cal_df %>% 
  drop_na(rest_type) %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Beverages and Desserts Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "beverages_dessert") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Convenience Stores

```{r}
final_cal_df %>% 
  filter(rest_type == "convenience") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Fast Food Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "fast food") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Sit-down Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "sit-down") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution by Food Category Overlayed

```{r}
final_cal_df %>% 
  drop_na(rest_type) %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution in Beverages and Desserts Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "beverages_dessert") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution in Convenience Stores

```{r}
final_cal_df %>% 
  filter(rest_type == "convenience") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution in Fast Food Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "fast food") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Trans Fat Distribution in Sit-down Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "sit-down") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution by Food Category Overlayed

```{r}
final_cal_df %>% 
  drop_na(rest_type) %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Beverages and Desserts Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "beverages_dessert") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Convenience Stores

```{r}
final_cal_df %>% 
  filter(rest_type == "convenience") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Fast Food Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "fast food") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Sit-down Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "sit-down") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```
